---
title: "HR Analysis and Vizualize with Highcharter"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

#Analysis
In this analysis, will investage various features associated to employees and the factors which causes attrition.
     Will analyse each factor one by one .Have used highcharter for visualization.
     

Loding the required libraries
```{r include=FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(highcharter)
library(readr)
library(tidyr)
library(RColorBrewer)

```

Get the data from csv file

```{r}

setwd("c:/testR/datasets/Kaggle/")
hrdata<-read.csv("Hr_comma_sep.csv")

summary(hrdata)
```

### Satisfaction Level
Let us compare the staisfaction level of the employess who had left along with the salary groupings

```{r echo=FALSE}
hchart(hrdata,"column", hcaes(x = left, y = satisfaction_level, group = salary,fill=salary))%>%hc_title(text = "Satisfaction Level Vs Salaray")


```
Distribution of Satisfaction level


```{r echo=FALSE}
left<-hrdata %>% filter(left==1)
notleft<-hrdata %>% filter(left==0)
hchart(density(left$satisfaction_level), area = TRUE,name="Left") %>%
  hc_add_series(density(notleft$satisfaction_level), area = TRUE,name="Not Left")
```
Employees with low satisfaction level has left the company, and people who still remain are having good satisfaction level.

check out - in which department,how many employees left 


```{r echo=FALSE}
 left<- hrdata %>% select(sales,satisfaction_level,left)%>%group_by(sales,left)%>%summarize(n_left=n())
hchart(left,"column",hcaes(x=sales,y=n_left,group=factor(left)))%>%hc_title(text="Departmentwise Attrition")
```
From the above graph, employee seems to high in sales, technical and support departments proportion to employees in that group.

Interesting to see that in Management department, attrition was very less compare to others.

### Number of Projects worked


```{r echo=FALSE}
p<-hrdata %>% group_by(number_project,left)%>% summarize(pn=n())

sp<-spread(p,left,pn)
colnames(sp)<-c("No of Projects","Not Left","Left")




hc1 <- highchart()%>% 
  hc_title(text = "Attrition Vs Number of Projects") %>% 
  hc_xAxis(categories = sp$`No of Projects`) %>% 
  hc_add_series(name = "Not Left", data = sp$`Not Left`,
               dataLabels = list(enabled = TRUE)) %>%
  hc_add_series(name = "Left", data = sp$Left,
                dataLabels = list(enabled = TRUE)) %>%
  hc_yAxis(title = list(text = "Employees") ) %>%
  hc_add_theme(hc_theme_sandsignika())

hc1


```
Employees who worked on more number of projects(>6) had left the company. Frequent change of projects might have impacted their satisfaction level.

###Time Spend in Company


```{r echo=FALSE}
df<-data.frame(table(hrdata$time_spend_company,hrdata$left))


df<-spread(df,Var2,Freq)
colnames(df)<-c("time_spent_company","Not Left","Left")




hc3 <- highchart()%>% 
  hc_title(text = "Attrition Vs Time Spent in Company") %>% 
  hc_xAxis(categories = df$time_spent_company) %>% 
  hc_add_series(name = "Not Left", data = df$`Not Left`,
                dataLabels = list(enabled = TRUE)) %>%
  hc_add_series(name = "Left", data = df$Left,
                dataLabels = list(enabled = TRUE)) %>%
  hc_yAxis(title = list(text = "Employees") ) %>%
  hc_add_theme(hc_theme_monokai())

hc3

```
Employees with average time spent has left the company, employees who had spent more didn't left.

###Promotion



Check out which salary grouped(Low ,Medium,High) got promoted in last 5 years


```{r echo=FALSE}
prom<-data.frame(table(prom=factor(hrdata$promotion_last_5years),hrdata$salary))


hc_p<-hchart(prom, "column",hcaes(x = Var2, y = Freq, group = prom)) %>%
  hc_plotOptions(column = list(stacking = "percent"))

hc_p<-hc_p %>%hc_title(text = "Percentage of Employess Promoted") %>% 
  hc_xAxis(title = list(text = "Salary")) %>% 
  hc_yAxis(title = list(text = "Promotion Percentage")) %>% 
  hc_chart(zoomType = "xy") 

hc_p
```

### Last Evaluation



```{r echo=FALSE}
hre<-data.frame(table(hrdata$last_evaluation,hrdata$left))

hc_p1<-hchart(hre, "scatter",hcaes(x = Var1, y = Freq,group=Var2))
  hc_p1<-hc_p1 %>%hc_title(text = "Last Evaluation Vs Attrition") %>% 
  hc_xAxis(title = list(text = "Last Evaluation")) %>% 
  hc_yAxis(title = list(text = "Employee Count")) %>% 
  hc_chart(zoomType = "xy") 

hc_p1




```
Last Evaluation score  in a treemap
```{r echo=FALSE}
heval<-data.frame(table(hrdata$last_evaluation))

df <- setNames(heval, c("name", "value"))
ds <- list_parse(df)
highchart() %>% 
  hc_title(text = "Last Evaluation Score vs Employees count") %>% 
  hc_add_series(data = ds, type = "treemap", colorByPoint = TRUE) 
```


Most of the Employees who left has got very low score in last evaluation, which is one of the important factor

Employees who had got very low evaluation score , but worked more than average monthly working hours.

```{r echo=FALSE}

hd<-hrdata %>% select(average_montly_hours,last_evaluation,left,time_spend_company)%>% filter(last_evaluation<0.5 & average_montly_hours>240)

highchart() %>% 
  hc_title(text = "Employees who got very low score") %>% 
  hc_add_series_scatter(hd$average_montly_hours,hd$last_evaluation,
                       hd$time_spend_company,hd$left)


```

###Salary
```{r echo=FALSE}
hchart(as.character(hrdata$salary), type = "pie") %>%hc_title(text = "Employess in Salary Group")

```
###Employees
In this company there exists several departments, lets find out which department has more employees
```{r echo=FALSE}

  
  hrsales<-data.frame(table(hrdata$sales))%>% arrange(desc(Freq))
hcbar <- highchart() %>% 
  hc_xAxis(categories = hrsales$Var1,title="Department") %>% 
  hc_yAxis(title = NULL) %>% 
  hc_add_series(data = hrsales$Freq, type = "bar", showInLegend = FALSE,
                name ="Employees",color = c("#2FB78B"))%>%
  hc_title(text="Employees Deparmentwise")

hcbar

```
###Work Accident
Will accident in work place has an impact over attrition, lets find out.

```{r echo=FALSE}
work<-hrdata %>% select(Work_accident,left,sales)%>% filter(Work_accident==1 )%>% group_by(sales,left)%>%summarize(wn=n())%>%arrange(desc(wn))

ggplot(work,aes(x=sales,y=wn,fill=factor(left)))+geom_bar(stat="identity",position="Dodge")+theme(legend.position = "bottom",axis.text.x = element_text(angle=90,hjust=0.2))+labs(title="Employees who left after Work Accident"
,y="count")+scale_fill_manual(values=c("#FFF98B","#FF9067"))
```

###Correlation

Finding out the correlation between the varaibles

```{r echo=FALSE}
  hchart.cor <- function(object, ...) {
    
    df <- as.data.frame(object)
    is.num <- sapply(df, is.numeric)
    df[is.num] <- lapply(df[is.num], round, 2)
    dist <- NULL
    
    x <- y <- names(df)
    
    df <- tbl_df(cbind(x = y, df)) %>% 
      gather(y, dist, -x) %>% 
      mutate(x = as.character(x),
             y = as.character(y)) %>% 
      left_join(data_frame(x = y,
                           xid = seq(length(y)) - 1), by = "x") %>% 
      left_join(data_frame(y = y,
                           yid = seq(length(y)) - 1), by = "y")
    
    ds <- df %>% 
      select_("xid", "yid", "dist") %>% 
      list_parse2()
    
    fntltp <- JS("function(){
                 return this.series.xAxis.categories[this.point.x] + ' ~ ' +
                 this.series.yAxis.categories[this.point.y] + ': <b>' +
                 Highcharts.numberFormat(this.point.value, 2)+'</b>';
                 ; }")
    cor_colr <- list( list(0, '#FF5733'),
                      list(0.5, '#F8F5F5'),
                      list(1, '#2E86C1')
    )
    highchart() %>% 
      hc_chart(type = "heatmap") %>% 
      hc_xAxis(categories = y, title = NULL) %>% 
      hc_yAxis(categories = y, title = NULL) %>% 
      hc_add_series(data = ds) %>% 
      hc_plotOptions(
        series = list(
          boderWidth = 0,
          dataLabels = list(enabled = TRUE)
        )) %>% 
      hc_tooltip(formatter = fntltp) %>% 
      hc_legend(align = "right", layout = "vertical",
                margin = 0, verticalAlign = "top",
                y = 25, symbolHeight = 280) %>% 
      hc_colorAxis(  stops= cor_colr,min=-1,max=1)
    }

hrdata2 <- hrdata[,1:8]
x <- cor(hrdata2)
hchart.cor(x)

```

Conclusion :

    As per the analysis, employess had left the company for following reasons
    1)Frequent change in projects, which has made the satisfactory level very low
    2)promotion and last evaluation are other important factors
    